---
title: "virus testing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
library(ggthemes)
library(paletteer)
library(ggplot2)
library(RColorBrewer)
```

#setting up aesthetics for later
```{r}
colors<- LaCroixColoR::lacroix_palette("Pamplemousse", type = "discrete")
```

#importing data
```{r}
viruses <- readr::read_csv("data/viruses.csv")
```

```{r}
viruses <- clean_names(viruses)
```

#cleaning data
```{r}
viruses <- viruses %>% 
  filter(level == "Complete") %>% 
  select(organism_name, organism_groups, level, size_mb, gc_percent, host, genes)
```

```{r}
viruses<-viruses %>% 
  separate(organism_groups, into = c("domain", "group", "family"),sep=";")
```

```{r}
viruses_mic <- viruses %>% 
  select(!domain, !group) %>% 
  filter(family != "unclassified", family != "Other")
```

#glimpse of data
```{r}
glimpse(viruses_mic)
```

#average genome size for all classified viruses
#clearly, mimiviruses stand out
#lots of variation
```{r}
viruses_mic %>% 
  group_by(family) %>% 
  summarize(size=mean(size_mb, na.rm = T)) %>% 
  ggplot(aes(x = family, y = size))+
  geom_col()+
  theme(axis.text.x = element_text(angle = 90, hjust=1))
```

#genome size versus common-ness
#inverse correlation?
```{r}
families_by_size <- viruses_mic %>% 
  group_by(family) %>% 
  summarize(size = mean(size_mb, na.rm=T)) %>% 
  arrange(desc(size))

families_by_size
```

```{r}
families_by_number <- viruses_mic %>% 
  group_by(family) %>% 
  summarize(num = n_distinct(organism_name)) %>% 
  arrange(desc(num))

families_by_number
```

```{r}
families_by <- full_join(families_by_number, families_by_size)
families_by
```

```{r}
families_by %>% 
  ggplot(aes(x=num, y=size))+
  geom_point()
```

#comparing overall genome size to number of genes
```{r}
viruses %>% 
  ggplot(aes(x=size_mb, y=genes))+
  geom_point()+
  geom_smooth()

```

```{r}
viruses %>% 
  filter(size_mb >= 1.0, genes <= 250)
```
```{r}
viruses %>% 
  filter(family=="Mimiviridae")
```
```{r}
viruses %>% 
  filter(genes != 0) %>% 
  ggplot(aes(x=size_mb, y=genes))+
  geom_point()+
  geom_smooth()
```

#comparing viruses with bacterial vs archaeal hosts
#bacteria: larger genome size, more genes, higher gc content
#similar gene to genome size ratio
```{r}
arch_bact <- viruses %>% 
  filter(host == "archaea"|host == "bacteria")
```

```{r}
colors <- LaCroixColoR::lacroix_palette("Pamplemousse", type = "discrete")
```

```{r}
display.brewer.pal(2,"PRGn")
```


```{r}
arch_bact %>% 
  group_by(host) %>% 
  summarize(size = mean(size_mb, na.rm = T)) %>%
  ggplot(aes(x=host, y=size))+
  geom_col()+scale_fill_brewer(palette = "Dark2")
```


```{r}
arch_bact %>% 
  group_by(host) %>% 
  summarize(size = mean(size_mb, na.rm = T)) %>%
  ggplot(aes(x=host, y=size))+
  geom_col()+
  #theme_solarized()+
  scale_fill_manual(values=colors)+
  labs(title = "Genome Size of Viruses by Host",
       x="Host",
       y="Genome Size (Mb)")
```

```{r}
arch_bact %>% 
  group_by(host) %>% 
  summarize(gc = mean(gc_percent, na.rm = T)) %>%
  ggplot(aes(x=host, y=gc))+
  geom_col()
```

```{r}
arch_bact %>% 
  group_by(host) %>% 
  summarize(gene = mean(genes, na.rm = T)) %>%
  ggplot(aes(x=host, y=gene))+
  geom_col()
```
```{r}
arch_bact %>% 
  filter(genes != 0) %>% 
  group_by(host) %>% 
  summarize(gene = mean(genes, na.rm = T)) %>%
  ggplot(aes(x=host, y=gene))+
  geom_col()
```
```{r}
arch_bact %>% 
  filter(genes != 0) %>% 
  mutate(gene_mb_ratio = genes/size_mb) %>% 
  group_by(host) %>% 
  summarize(gene_content = mean(gene_mb_ratio, na.rm = T)) %>%
  ggplot(aes(x=host, y=gene_content))+
  geom_col()
```
 
#satellites vs non satellites
```{r}
sat_other <- viruses %>% 
  filter(group == "Satellites" | group == "Other")
```

```{r}
sat_other %>% 
  group_by(group) %>% 
  summarize(size = mean(size_mb)) %>% 
  ggplot(aes(x=group, y=size)) +
  geom_col()
```

```{r}
sat_other %>% 
  group_by(group) %>% 
  summarize(gene = mean(genes)) %>% 
  ggplot(aes(x=group, y=gene)) +
  geom_col()
```

```{r}
sat_other %>% 
#  filter(genes != 0) %>% 
  mutate(gene_mb_ratio = genes/size_mb) %>% 
  group_by(group) %>% 
  summarize(gene_content = mean(gene_mb_ratio, na.rm = T)) %>%
  ggplot(aes(x=group, y=gene_content))+
  geom_col()
```

