---
title: "Midterm 1"
author: "Claire Fargeix"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your code should be organized, clean, and run free from errors. Be sure to **add your name** to the author header above. You may use any resources to answer these questions (including each other), but you may not post questions to Open Stacks or external help sites. There are 12 total questions.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

This exam is due by **12:00p on Thursday, January 28**.  

## Load the tidyverse
If you plan to use any other libraries to complete this assignment then you should load them here.
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## Questions
**1. (2 points) Briefly explain how R, RStudio, and GitHub work together to make work flows in data science transparent and repeatable. What is the advantage of using RMarkdown in this context?**  
R is a coding language that programmers can use via the GUI RStudio. GitHub allows for programmers to upload and share their code with people around the world. RMarkdown is useful in this context because it provides an organized and easy to read document so data scientist can understand the code that is being shared.

**2. (2 points) What are the three types of `data structures` that we have discussed? Why are we using data frames for BIS 15L?**  
We have discussed vectors, data matrices, and data frames. We are using data frames because they are an efficient way of organizing a lot of data in a spreadsheet like form.  

In the midterm 1 folder there is a second folder called `data`. Inside the `data` folder, there is a .csv file called `ElephantsMF`. These data are from Phyllis Lee, Stirling University, and are related to Lee, P., et al. (2013), "Enduring consequences of early experiences: 40-year effects on survival and success among African elephants (Loxodonta africana)," Biology Letters, 9: 20130011. [kaggle](https://www.kaggle.com/mostafaelseidy/elephantsmf).  

**3. (2 points) Please load these data as a new object called `elephants`. Use the function(s) of your choice to get an idea of the structure of the data. Be sure to show the class of each variable.**
```{r}
elephants <- readr::read_csv("data/ElephantsMF.csv")
```
```{r}
glimpse(elephants)
```

**4. (2 points) Change the names of the variables to lower case and change the class of the variable `sex` to a factor.**
```{r}
elephants <- elephants %>%
  rename(age=Age, height=Height, sex=Sex)
```
```{r}
elephants$sex <- as.factor(elephants$sex)
```
```{r}
glimpse(elephants)
```

**5. (2 points) How many male and female elephants are represented in the data?**  
There are 138 male elephants and 150 female elephants.
```{r}
anyNA(elephants$sex)
```

```{r}
elephants %>% 
  count(sex=="F")
```
**6. (2 points) What is the average age all elephants in the data?**  
The average age of all the elephants in the data is 10.97.
```{r}
elephants %>% 
  summarise(mean_age = mean(age))
```

**7. (2 points) How does the average age and height of elephants compare by sex?**  
Both the average age and the average height of the elephants is higher for the female elephants.
```{r}
elephants %>% 
  group_by(sex) %>% 
  summarise(mean_age = mean(age), mean_height = mean(height))
```

**8. (2 points) How does the average height of elephants compare by sex for individuals over 25 years old. Include the min and max height as well as the number of individuals in the sample as part of your analysis.**  
Male elephants over 25 years old are on average taller than the female elephants over 25 years of age.
```{r}
elephants %>% 
  filter(age >= 25) %>% 
  group_by(sex) %>% 
  summarize(mean_height = mean(height),
            min_height = min(height),
            max_height = max(height),
            n = n())
```

For the next series of questions, we will use data from a study on vertebrate community composition and impacts from defaunation in [Gabon, Africa](https://en.wikipedia.org/wiki/Gabon). One thing to notice is that the data include 24 separate transects. Each transect represents a path through different forest management areas.  

Reference: Koerner SE, Poulsen JR, Blanchard EJ, Okouyi J, Clark CJ. Vertebrate community composition and diversity declines along a defaunation gradient radiating from rural villages in Gabon. _Journal of Applied Ecology_. 2016. This paper, along with a description of the variables is included inside the midterm 1 folder.  

**9. (2 points) Load `IvindoData_DryadVersion.csv` and use the function(s) of your choice to get an idea of the overall structure. Change the variables `HuntCat` and `LandUse` to factors.**
```{r}
vertebrates <- readr::read_csv("data/IvindoData_DryadVersion.csv")
```
```{r}
glimpse(vertebrates)
```

```{r}
vertebrates$HuntCat <- as.factor(vertebrates$HuntCat)
vertebrates$LandUse <- as.factor(vertebrates$LandUse)
```

**10. (4 points) For the transects with high and moderate hunting intensity, how does the average diversity of birds and mammals compare?**  
At both moderate and high hunting intensity, the average mammal diversity is slightly higher than the average bird diversity. Additionally, the diversities of both mammals and birds are higher at high hunting intensity than moderate intensity.
```{r}
vertebrates %>%
  filter(HuntCat=="High" | HuntCat=="Moderate") %>% 
  group_by(HuntCat) %>%
  summarize(avg_mammal_diversity = mean(Diversity_MammalSpecies),
            avg_bird_diversity = mean(Diversity_BirdSpecies))
```

**11. (4 points) One of the conclusions in the study is that the relative abundance of animals drops off the closer you get to a village. Let's try to reconstruct this (without the statistics). How does the relative abundance (RA) of apes, birds, elephants, monkeys, rodents, and ungulates compare between sites that are less than 5km from a village to sites that are greater than 20km from a village? The variable `Distance` measures the distance of the transect from the nearest village. Hint: try using the `across` operator.**  
The relative abundance of animals is lower when less than 5km from villages across all of animals except birds and rodents. There is a slight increase for rodents and a large increase for birds when going closer to villages.
```{r}
vertebrates %>%
  filter(Distance >= 20) %>%
  summarize(across(contains("RA_"), mean))
```
```{r}
vertebrates %>%
  filter(Distance <= 5) %>%
  summarize(across(contains("RA_"), mean))
```

**12. (4 points) Based on your interest, do one exploratory analysis on the `gabon` data of your choice. This analysis needs to include a minimum of two functions in `dplyr.`**  
```{r}
vertebrates %>% 
  group_by(LandUse) %>% 
  summarize(across(contains("Veg_"), mean))
```

